### 2012-11-30 doc/mandc2013/theory_paper/slattery_wilson_evans.tex
updating figures and adding gnumeric data file

In a domain decomposed situation, not all histories will remain within
the domain they started in and must instead be communicated. This
communication, expected to be expensive, was analyzed by Siegel and
colleagues for idealized, load balanced situations for full nuclear
reactor core Monte Carlo simulations \cite{siegel_analysis_2012}.  To
quantify the number of particles that leak out of the local domain
they define a leakage fraction, $\Lambda$, as:
\begin{equation}
  \Lambda = \frac{average\ \#\ of\ particles\ leaving\ local\ domain}
          {total\ of\ \#\ of\ particles\ starting\ in\ local\ domain}\:.
          \label{eq:leakage_fraction}
\end{equation}
For their studies, they assumed that the value of $\Lambda$ was
dependent on the total cross section of the system via the Wigner
rational approximation. Outlined more thoroughly by Hwang's chapter in
\cite{azmy_nuclear_2010}, we will use both the Wigner rational
approximation and the mean chord approximation as a means to estimate
the leakage fraction.

@@

To compute the leakage fraction numerically, \sn{3}{5} histories were
sampled from a uniform source of strength unity over the global
domain. At the start of a stage of histories, the number of histories
starting in the domain was computed and as the stage progressed, the
number of histories that exited the domain was tallied with the ratio
of the two numbers providing a numerical measure for the leakage
fraction. Figure~\ref{fig:measured_length} gives the domain leakage
measurements for the domain in the center of the global grid as well
as the analytic result computed by
Eqs~(\ref{eq:wigner_domain_leakage}) and
(\ref{eq:mean_chord_domain_leakage}) as a function of the iteration
matrix spectral radius.
\begin{figure}[ht!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{leakage_variation.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage as a function
        of the iteration matrix spectral radius.} \textit{To test the
        behavior with respect to domain size, $n_i=50$, $n_i=100$,and
        $n_i=200$ were used. The red data was computed numerically by a
        domain-decomposed adjoint Neumann-Ulam implementation, the black
        fine-dashed data was generated by
        Eq~(\ref{eq:mean_chord_domain_leakage}) using the mean-chord
        approximation, and the dashed-dotted black data was generated by
        Eq~(\ref{eq:wigner_domain_leakage}) using the Wigner rational
        approximation.}}
    \label{fig:measured_leakage}
  \end{spacing}
\end{figure}
Again, we note good qualitative agreement between the measured and
analytic quantities but we begin to see the limits of the leakage
approximations. To compare the quality of the two approximations, the
absolute error between the computed leakage fraction and that
generated by the Wigner rational and mean chord approximations is
plotted in Figure~\ref{fig:leakage_error} for all domain sizes
tested. 
\begin{figure}[ht!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{leakage_error.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage absolute error
        as a function of the iteration matrix spectral radius.}
      \textit{To test the behavior with respect to domain size, $n_i=50$
        (blue), $n_i=100$ (red), and $n_i=200$ (yellow) were used. The
        dashed lines represent the error using the Wigner rational
        approximation while the solid lines represent the error using
        the mean-chord approximation.}}
    \label{fig:leakage_error}
  \end{spacing}
\end{figure}
From these error results, the mean chord approximation is
shown to have a lower error for ill-conditioned systems as compared
to the Wigner approximation while the Wigner approximation produces
less error for more well-conditioned systems. In general, the mean
chord approximation is a better choice to estimate the leakage
fraction in a domain from the adjoint Neumann-Ulam method and except
for a single data point with $n_i=50$, the mean chord approximation
yielded leakage fractions within 0.05 of the measured results.

>>>>

In a domain decomposed situation, not all histories will remain within
the domain they started in and must instead be communicated. This
communication, expected to be expensive, was analyzed by Siegel and
colleagues for idealized, load balanced situations for full nuclear
reactor core Monte Carlo simulations \cite{siegel_analysis_2012}.  To
quantify the number of particles that leak out of the local domain
they define a leakage fraction, $\Lambda$, as:
\begin{equation}
  \Lambda = \frac{average\ \#\ of\ particles\ leaving\ local\ domain}
          {total\ of\ \#\ of\ particles\ starting\ in\ local\ domain}\:.
          \label{eq:leakage_fraction}
\end{equation}
For their studies, Siegel and colleagues assumed that the value of
$\Lambda$ was dependent on the total cross section of the system via
the Wigner rational approximation. Outlined more thoroughly by Hwang's
chapter in \cite{azmy_nuclear_2010}, we will use both the Wigner
rational approximation and the mean chord approximation as a means to
estimate the leakage fraction.

@@

To compute the leakage fraction numerically, \sn{3}{5} histories were
sampled from a uniform source of strength unity over the global
domain. At the start of a stage of histories, the number of histories
starting in the domain was computed and as the stage progressed, the
number of histories that exited the domain was tallied with the ratio
of the two numbers providing a numerical measure for the leakage
fraction. Figure~\ref{fig:measured_length} gives the domain leakage
measurements for the domain in the center of the global grid as well
as the analytic result computed by
Eqs~(\ref{eq:wigner_domain_leakage}) and
(\ref{eq:mean_chord_domain_leakage}) as a function of the iteration
matrix spectral radius.
\begin{figure}[ht!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{leakage_variation.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage as a
        function of the iteration matrix spectral radius.} \textit{To
        test the behavior with respect to domain size, $n_i=50$,
        $n_i=100$,and $n_i=200$ were used. The red data was computed
        numerically by a domain-decomposed adjoint Neumann-Ulam
        implementation, the black dashed data was generated by
        Eq~(\ref{eq:mean_chord_domain_leakage}) using the mean-chord
        approximation, and the dashed-dotted black data was generated
        by Eq~(\ref{eq:wigner_domain_leakage}) using the Wigner
        rational approximation.}}
    \label{fig:measured_leakage}
  \end{spacing}
\end{figure}
Again, we note good qualitative agreement between the measured and
analytic quantities but we begin to see the limits of the leakage
approximations. To compare the quality of the two approximations, the
absolute error between the computed leakage fraction and that
generated by the Wigner rational and mean chord approximations is
plotted in Figure~\ref{fig:leakage_error} for all domain sizes
tested. 
\begin{figure}[ht!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{leakage_error.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage absolute error
        as a function of the iteration matrix spectral radius.}
      \textit{To test the behavior with respect to domain size, $n_i=50$
        (green), $n_i=100$ (blue), and $n_i=200$ (red) were used. The
        dashed lines represent the error using the Wigner rational
        approximation while the solid lines represent the error using
        the mean-chord approximation.}}
    \label{fig:leakage_error}
  \end{spacing}
\end{figure}
From these error results, the mean chord approximation is
shown to have a lower error for ill-conditioned systems as compared
to the Wigner approximation while the Wigner approximation produces
less error for more well-conditioned systems. In general, the mean
chord approximation is a better choice to estimate the leakage
fraction in a domain from the adjoint Neumann-Ulam method and except
for a single data point with $n_i=50$, the mean chord approximation
yielded leakage fractions within 0.05 of the measured results.
